---
title: "Analysis for E:I ratio"
author: "Bart"
date: "10/29/2021"
output: html_document
---

#Primary analyses  


```{r setup, include=FALSE}
library(tidyverse)
library(sjPlot)
library(mgcv)
library(broom)
library(purrr)
require(ggplot2)
library(cowplot)
require(dplyr)
knitr::opts_chunk$set(fig.height = 6,fig.width = 6,echo = F,warning = FALSE)
# setwd("/cbica/projects/alpraz_EI/SubmissionGithub/")
source("Alpraz_viz_functions.R")
font_size <- 16
unimodal_color = "#1c405eff"
transmodal_color = "#285c27ff"
theme_set(theme_classic(base_family = "sans",base_size = font_size))
theme_replace(axis.text=element_text(colour = "black",size = font_size))

# Load all the data
alpraz_info = read.csv('input/alpraz_sublist_FD2.csv')
alpraz_demo = read.csv('input/Alpraz_demographics.csv')
alpraz_info <- alpraz_info %>% left_join(alpraz_demo,by=c("subid"="bblid", "sesid"="fmriid"))
# Sample characteristics
alpraz_sample <- alpraz_info%>%
  filter(!is.na(subid))
print(alpraz_sample %>% 
  select(subid,age.at.fmri1,sex_M0F1,motion_pass,exists) %>%
  filter(exists==1 & motion_pass==1) %>%
  unique()%>%
  summarize(n=n(),
            min_age = min(age.at.fmri1,na.rm=T),
            max_age = max(age.at.fmri1,na.rm=T),
            mean_age = mean(age.at.fmri1,na.rm=T),
            sd_age=sd(age.at.fmri1),
            males = sum(sex_M0F1==1),
            females = sum(sex_M0F1==0)))

PNCData <- read.table('input/PNC_df.csv',header = T,sep=",")
# Get exclusions due to poor data quality and health data (low threshhold normal used here)
FD_thresh <- .3 #threshold for mean FD
PNCData <- PNCData %>%
  mutate(FD_exclude = FD>FD_thresh)

PNCData$exclusions <-  PNCData$ltnExcludev2==1 | 
  PNCData$FD_exclude==1 |
  is.na(PNCData$idemoBehAllEmoNrCount)|
  PNCData$idemoExcludeVoxelwise==1

print(PNCData %>% 
  select(subid,ageAtCnb1,sex,exclusions) %>%
  filter(exclusions==0) %>%
  summarize(n=n(),
            min_age = min(ageAtCnb1/12,na.rm=T),
            max_age = max(ageAtCnb1/12,na.rm=T),
            mean_age = mean(ageAtCnb1/12,na.rm=T),
            sd_age=sd(ageAtCnb1/12),
            males = sum(sex==1),
            females = sum(sex==2)))


```


```{r functions,echo=FALSE}
# A number of functions that are used.

# This one just merges data.frames and deals with variable types, etc.
data_mgmt <- function(fname){
  pred_results <- readRDS(fname)%>%
    left_join(PNCData,by = "subid")
  pred_results$oPred <- ordered(pred_results$pred,levels=c(1,0),labels=c("Placebo","Drug")) #1 is placebo, 0 is drug.
  pred_results<-pred_results%>%
    mutate(age = ageAtCnb1/12,age2=age^2,
           oSex = ordered(sex,levels=c(1,2),labels=c("male","female")),
           oRace = ordered(race2),
           fRace = factor(race2),
           weights = 1/FD,
           distance = as.numeric(distance),
           nPred = as.numeric(pred==0)) #Note Drug is now 1 and placebo is now 0 (for ease of visualization)
  return(pred_results)
}

# This runs models and plots the outputs.
plot_results <- function(pred_results,title_text,plot_color="black"){
  
  # Fit distance from the classification hyperplane, a continuous measure, as a function of age and covariates using GAM.
  cat("\nDevelopmental models for  distance\n")
  pred_results$Distance = as.numeric(pred_results$distance)

  # GAM
  cat("\nGAM\n")
  gam_model <- gam(Distance ~s(age,k=4)+ oSex+FD+ idemoBehAllEmoNrCount, data = pred_results,
                 subset = exclusions==0)
  print(summary(gam_model))
  tab_model(gam_model)

  DistGAM <- visualize_model(gam_model,smooth_var = "age",int_var = NULL,group_var = NULL,derivative_plot = FALSE,line_color=plot_color)
  ggsave(sprintf('figs/%s_distance_gam.svg',title_text),DistGAM,device = 'svg',height = 5,width = 5)
}
```

## Schaefer 400
Classifier performance and age effects when using all the schaefer 400 regions.  

```{r 400, warning=FALSE}
pred_results <- data_mgmt('output/PNC_predictions/schaefer400x7_aal_all.rds')
display_results("schaefer400x7_aal",perm_results = T,)
plot_results(pred_results,"Schaefer400_All_Regions")
```

Explore the sessions that were mis-classified for Table S1.  

```{r misclassifications_data_org, include=FALSE}
# Load the classifier results
results <- readRDS("output/drug_classification/schaefer400x7_aal_GSR_all_svm_1_permute_on_results.rds")
pred_data <- results[[1]]$pred_data[[1]] 

# Since we fit the model 100 times with different CV folds, we take the mean decision value across all folds and use this to create an aggregate/overall prediction of drug vs placebo.
dec_folds <- data.table::rbindlist(results[[1]]$pred_data,idcol = "fold")%>%
  pivot_wider(names_from = "fold","values_from"="decisionValues",id_cols = c("subid","sesid")) #data organization
pred_data$mean_dec_vals <- rowMeans(dec_folds[,3:dim(dec_folds)[2]]) #calculate the mean across folds
pred_data <- pred_data %>% 
  mutate(overall_prediction = case_when(mean_dec_vals>0 ~ 0,mean_dec_vals<0~1),#get the overall prediction from the mean values.
         accurate_classification = as.numeric(overall_prediction == drug)) # check if it was accurately classified

# Now join with a data table of covariates
alpraz_covs = read.table('./input/Alpraz_covariates.csv',sep = ",",header = T)
alpraz_df <- pred_data %>% select(subid,sesid,accurate_classification,drug,mean_dec_vals) %>%
  left_join(alpraz_covs, by = c("subid"="bblid","sesid"="fmriid")) %>%
  left_join(alpraz_info %>% select(sesid,subid,FD), by = c("subid","sesid"))%>%
  mutate(Sex=factor(M0.F1),Condition = factor(drug),Order = factor(day),Age=age.at.fmri1)
  
```


```{r Table S1,include=TRUE}
# Conduct a mass univariate analysis to evaluate whether demographic or experimental procedures impact mis-classification.
## For continuous variables, use a t-test.
t.test.results <- alpraz_df %>%
  pivot_longer(cols=c(Age, Relaxed,alpraz.ng.mL., STAI.TRAIT, STAI.PREMED, STAI.PRESCAN,STAI.POSTSCAN,Prescan.Postscan, SISTOTAL,FD),
               values_to="value",
               names_to="covariate")%>%
  select(covariate,where(is.numeric))%>%
  group_by(covariate)%>%
  do(broom::tidy(t.test(value ~ accurate_classification, data = .)))%>%
  select(covariate,statistic,p.value)

## For factors, use a chi-square test.
chsq.results <- alpraz_df %>%
  select(Sex,Condition,Order) %>%
  summarise_all(funs(broom::tidy(chisq.test(.,alpraz_df$accurate_classification))))%>%
  pivot_longer(cols = everything())%>%
  unnest(cols = value)%>%
  rename(covariate=name)%>%
  select(covariate,statistic,p.value)

## Stick them together and print a table.
covariate.results <- rbind(t.test.results,chsq.results)

knitr::kable(covariate.results, digits = 2,
                caption = "Effect of covariates on accurate classification",
                col.names = c("Covariate","Statistic","p-value"))%>%
  kableExtra::kable_classic_2()
```

## Transmodality analyses
Looking at the classifier performance and age effects for models trained on connections to the most transmodal or unimodal regions only.  

### Transmodal

```{r transmodal}
pred_results <- data_mgmt('output/PNC_predictions/schaefer400x7_aal_transmodal25.rds')
alpraz_results <- readRDS("output/drug_classification/schaefer400x7_aal_GSR_transmodal25_svm_1_permute_on_results.rds")
display_results(atlasname = "schaefer400x7_aal",perm_results = TRUE,results = alpraz_results,subdivision = "transmodal25")
b <- alpraz_results[[1]]
cat("Results for classification:\n")
cat(sprintf("\n Number of features: %d\n",alpraz_results[[3]]))
cat("\nAccuracy:\n")
print(b$accuracy)
cat(sprintf("p = %1.3f\n",b$p.value))

pred_a_400 = pred_results
plot_results(pred_a_400,"Schaefer400_Transmodality",plot_color = transmodal_color)

```

### Unimodal

```{r unimodal}
pred_results <- data_mgmt('output/PNC_predictions/schaefer400x7_aal_unimodal25.rds')
alpraz_results <- readRDS("output/drug_classification/schaefer400x7_aal_GSR_unimodal25_svm_1_permute_on_results.rds")
display_results(atlasname = "schaefer400x7_aal",perm_results = T,results = alpraz_results,subdivision="unimodal25")
b <- alpraz_results[[1]]
cat("Results for classification:\n")
cat(sprintf("\n Number of features: %d\n",alpraz_results[[3]]))
cat("\nAccuracy:\n")
print(b$accuracy)
cat(sprintf("p = %1.3f\n",b$p.value))

pred_s_400 = pred_results
plot_results(pred_s_400,"Schaefer400_Unimodal",plot_color = unimodal_color)

```

### Transmodal vs. unimodal development    

Looking at the difference in the distance metric between the transmodal and the unimodal classification models. We want to test the hypothesis that transmodal areas are undergoing the most pronounced E:I development in this age range. So, we test whether the w/in subject difference in transmodal and unimodal distances increases with age.    

```{r sens_assoc diff,fig.height=8,fig.width=8}

pred_a_400 <- pred_a_400 %>%
  mutate(a_distance=distance,
         a_nPred = nPred,
         transmodal_classifier = a_distance)

pred_s_400 <- pred_s_400 %>%
  mutate(s_distance=distance,
         s_nPred = nPred,
         unimodal_classifier = s_distance)
pred_join_400 <- pred_a_400 %>%
  left_join(pred_s_400,by=c("subid","age","FD","oSex","idemoBehAllEmoNrCount"))%>%
  pivot_longer(cols = c(transmodal_classifier,unimodal_classifier),names_to = "subdivision",values_to = "Distance")
difference_df <- pred_a_400 %>%
  left_join(pred_s_400,by=c("subid","age","FD","oSex","idemoBehAllEmoNrCount","exclusions"))%>%
  filter(exclusions==0)

# Calculate residualized change in distance between the transmodal and unimodal models
res_mod <- mgcv::gam(a_distance ~ s(s_distance),data = difference_df)
difference_df$res_change <- residuals(res_mod)

# Fit the residualized change as a function of age (with covariates)
gmod_diff <-mgcv::gam(res_change ~ oSex+FD+idemoBehAllEmoNrCount+ s(age,k=4),
                  data = difference_df)
summary(gmod_diff)
res_change_plot <- visualize_model(modobj = gmod_diff,smooth_var = "age",int_var = NULL,group_var = NULL,derivative_plot = F)
ggsave("figs/res_change_plot.svg",res_change_plot,device = "svg", height = 5,width = 5,units = "in")


```

## Clinical effects

Now we test whether the development of E:I balance is moderated by dimensions of psychopathology. Specifically, we extracting the clinical factor scores from the GOASSESS data and looking for age-by-factor-score interactions.  As psychopathology emerges during adolescence, the hypothesis is that greater symptom load will be impact developmental trajectories of E:I.  

```{r clinical, eval=TRUE,include=TRUE,fig.height=2,fig.width=3}
clinical <- read.csv('input/n1601_goassess_itemwise_bifactor_scores_20161219.csv')
clinical$subid=clinical$bblid

pred_results <- pred_a_400
pred_clinical <- pred_results %>%
  left_join(clinical,by = "subid")%>%
  mutate(distance = as.numeric(distance),abs_distance = abs(distance))

pred_clinical$exclusions <-  pred_clinical$healthExcludev2==1 | 
  pred_clinical$FD_exclude==1 |
  pred_clinical$idemoExcludeVoxelwise==1 

print(pred_clinical %>% 
  select(subid,age,sex,exclusions) %>%
  filter(exclusions==0) %>%
  summarize(n=n(),
            min_age = min(age,na.rm=T),
            max_age = max(age,na.rm=T),
            mean_age = mean(age,na.rm=T),
            sd_age=sd(age),
            males = sum(sex==1),
            females = sum(sex==2)))

# Run the model
gam_model <- gam(distance ~ ti(age,k=4)+ oSex+FD + idemoBehAllEmoNrCount + 
               ti(overall_psychopathology_4factorv2,k=4) + 
               ti(psychosis_4factorv2,k=4) + 
               ti(mood_4factorv2,k=4) + 
               ti(externalizing_4factorv2,k=4) + 
               ti(phobias_4factorv2,k=4) +
               ti(age,overall_psychopathology_4factorv2,k=4)+ 
               ti(age,psychosis_4factorv2,k=4)+
               ti(age,phobias_4factorv2,k=4)+
               ti(age,externalizing_4factorv2,k=4)+
               ti(age,mood_4factorv2,k=4),
             data = pred_clinical,subset = exclusions==0)
summary(gam_model)
visualize_model(gam_model,smooth_var = "age",int_var =  "mood_4factorv2",group_var = NULL)

```